---
title: "Wang_Code"
author: "Claire Willing"
date: "2024-11-20"
output: html_document
---
Read in libraries
```{r warning=FALSE}
# Function to install and load packages
install_and_load <- function(packages) {
  for (pkg in packages) {
    if (!requireNamespace(pkg, quietly = TRUE)) {
      if (pkg == "phyloseq") {
        if (!requireNamespace("BiocManager", quietly = TRUE)) {
          install.packages("BiocManager")
        }
        BiocManager::install(pkg)
      } else {
        install.packages(pkg)
      }
    }
    library(pkg, character.only = TRUE)
  }
}

# List of required packages
required_packages <- c(
  "phyloseq",  # For working with sequencing data
  "dplyr",     # For data manipulation
  "ggplot2",   # For plotting
  "readxl",    # For reading Excel files
  "writexl",   # For writing Excel files
  "tibble",    # For converting columns to row names
  "stringr",   # For string manipulation
  "viridis",    # For color palettes
  "ggeffects",
  "broom"
)

# Install and load packages
install_and_load(required_packages)
```
Get working directory
```{r}
working_directory <- getwd()
#unix (linux,mac)
its_phyloseq <- paste(working_directory, "WangPhyloseqData.rds", sep="/")
# Windows
#its_phyloseq <- paste(working_directory, "FischerPhyloseqData.rds", sep="\")
```

Pull in the file
```{r setup, include=FALSE}
Data_Raw <- readRDS(its_phyloseq)
```

Rarification
```{r message=FALSE, warning=FALSE}
set.seed(5)
RARE=rarefy_even_depth(Data_Raw, sample.size = 5000)
save(RARE, file = "RARE.rda")
load(file = "RARE.rda")

outlier_samples <- c("CW2","CW18", "CW4") 
RARE <- prune_samples(!sample_names(RARE) %in% outlier_samples, RARE)
```

Ordination (NMDS) using default of bray for distance (bray is default method for ordinate function)
```{r message=FALSE, warning=FALSE}
#make sure the variables are factors
sample_data(RARE)$block <- as.factor(sample_data(RARE)$block)
sample_data(RARE)$logdistance <- as.numeric(sample_data(RARE)$logdistance)

ORDtotalvals<-ordinate(RARE, "NMDS")
ORD<-plot_ordination(RARE, ORDtotalvals, color="logdistance",shape="block")+
  theme_minimal()
ORD

```
Barplots
```{r}
#merge by grouping
ps_merged <- merge_samples(RARE, "distance")
#turn into relative abundance
ps_rel <- transform_sample_counts(ps_merged, function(x) x / sum(x))

#aggregate at taxonomic level of interest (this part is slow)
ps_taxa <- tax_glom(ps_rel, taxrank = "Phylum")

#melt for data visualization
ps_melt <- psmelt(ps_taxa)
ps_melt$order[ps_melt$Abundance < 0.03] <- "Other"
ps_melt <- ps_melt %>%
  arrange(logdistance) %>%
  mutate(Sample = factor(Sample, levels = unique(Sample)))

barplot<-ggplot(ps_melt, aes(x = Sample, y = Abundance, fill = Phylum)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(x = "Distance from forest edge (m)", y = "Relative Abundance", fill = "Phylum") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  scale_fill_viridis_d(option = "plasma")

barplot
```

#Question 1: What is the relationship between total nitrogen content of the substrate and the dispersal distance of Agaricales, Cantharellales, and Russulales?

##Hypothesis 1: Higher amounts of nitrogen will have a positive correlation with dispersal distance. This study uses dispersal distance as a proxy for spore density; access to more nutrients will positively affect spore production/dispersal.

###H1 Statistics
```{r}
##for simplicity, I just took the entire Class of Agaricomycetes; this included Russulales, Cantheralles, Polyporales, Hymenochaetales, and Agaricales

Agaricomycetes<-subset_taxa(RARE, Class=="Agaricomycetes")
Agaricomycetes_df <- psmelt(Agaricomycetes)

# Assuming nitrogen is a column in sample data and 'dispersal_distance' is also available
Agaricomycetes_df$TN <- sample_data(Agaricomycetes_df)$TN  
Agaricomycetes_df$logdistance <- sample_data(Agaricomycetes_df)$logdistance  

N_Dispersal_model <- lm(logdistance ~ TN, data = Agaricomycetes_df)

summary(N_Dispersal_model)

table_N <- tidy(N_Dispersal_model)
View(table_N)
```
###H1 Plot
```{r}
# Generate predictions for mean relative humidity
predictions_nitrogen <- ggpredict(N_Dispersal_model, terms = c("TN"))

# Plot predictions for mean relative humidity
plot_N_distance <- ggplot(predictions_nitrogen, aes(x = x, y = predicted)) +
  geom_line(size = 1, color = "blue") +  # Prediction line
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "blue", alpha = 0.2) +  # Confidence intervals
  theme_minimal() +
  labs(
    x = "Total Nitrogen",
    y = "Predicted Dispersal Distance (Agaricomyetes)"
  )
plot_N_distance
```
#Question 2: Does species diversity decrease across distance?
##Hypothesis 2: Since most fungal species are unable to distribute spores across long distances, we predict that species diversity will decrease as distance increases.

###H2 Plots
```{r}
richness_df <- estimate_richness(RARE, measures = c("Observed", "Shannon"))
richness_df$Sample<-rownames(richness_df)
richness_df <- richness_df %>%
mutate(Sample = gsub("[.-]", "-", Sample))

SampleData<-as(sample_data(RARE), "data.frame")
SampleData$Sample<-rownames(SampleData)

richness_df$Observed <- as.numeric(richness_df$Observed)
richness_df$Shannon <- as.numeric(richness_df$Shannon)


Richnessdf<-full_join(richness_df, SampleData)

richness_plot <- ggplot(Richnessdf, aes(x = logdistance, y = Shannon)) +
geom_point(color = "darkgreen", alpha = 0.7) +
geom_smooth(method = "lm", se = FALSE, color = "darkgray") +
theme_minimal() +
theme(
panel.grid.major = element_line(color = "gray85"),
panel.grid.minor = element_blank(),
panel.background = element_rect(fill = "white", color = "gray90")) +
labs(
x = "Log of Distance (log distance)",
y = "Shannon Diversity Index (H')")

richness_plot
```
###H2 statistics
```{r}
lm_model <- lm(Shannon ~ logdistance, data = Richnessdf)
summary(lm_model)  

table <- tidy(lm_model)
View(table)
```

View(RARE@sam_data)
